---
title: "DSST 389: Final Project"
author: "Anna Phillips"
date: today
format: pdf
editor: visual
df-print: kable
---

# Introduction: Stephen Sondheim

Stephen Sondheim is one of Broadway's most beloved and successful lyricists/composers. He is a Tony Award, Academy Award, and Grammy Award winner, with a corpus of work ranging from Broadway musicals to television and movie scores. Sondheim's shows are known for their exploration of relationships, morality, and love -- heavily facilitated by Sondheim's lyrical verbosity and musical complexity. In an [interview](https://www.nytimes.com/2000/03/12/magazine/conversations-with-sondheim.html) with *The New York Times Magazine*, Sondheim reflected that "[he had] always conscientiously tried not to do the same thing twice," which was no easy feat when competing in a Broadway market that had turned "corporate" and "cookie-cutter."[^1] Because of this, even though Sondheim grapples with similar themes in his shows, his characters and narratives feel brand new every time, ranging from homicidal barbers in the 18th century, to fairytale characters lost together in the forest, to musings on marriage in affluent New York social circles, to the complex inner world of 19th-century artists and their struggles to devote time outside of their work. 

[^1]: Rich, Frank. 2000. "Conversations with Sondheim." *The New York Times Magazine*. https://www.nytimes.com/2000/03/12/magazine/conversations-with-sondheim.html

## So what?

Sondheim's shows grapple with life's hard questions, and his lyrics often trend toward the existential,[^2] allowing the audience to sympathize with broad themes -- despite the often hyper-specific nature of Sondheim's shows. Interestingly, through initial data exploration, I identified 'love' as the most frequent term in the corpus as a whole, and I will use multiple textual analysis methods to explore the significance of 'love' in Sondheim's corpus. First, I will use tf-idf to explore important terms for each show, and then I will use sentiment analysis to provide more context for the shows and their relationship to each other before moving to topic modeling. With topic modeling, I will return to the topic of 'love' and explore important terms across the entire corpus, as well as between the contrasting individual shows. Sondheim is known for exploring dark themes and subject matter, so it may be surprising that love is the most prevalent topic -- but when you explore his lyrics, you'll find that Sondheim explores multiple facets of love in his work (which includes darkness, revenge, and murder), and this complexity is what makes him so compelling as a lyricist.

[^2]: Kohn, Alfie. 2000. "The Case for Sondheim as Existentialist." *The Sondheim Review*. https://www.alfiekohn.org/article/case-sondheim-existentialist/

## Data Description

In order to perform my planned textual analysis, I had to create my own corpus of Sondheim musicals. To do this, I manually copy-and-pasted lines from eight shows into their own separate data sets and imported them to R. In terms of criteria, I chose his most famous shows, in combination with shows that have won the Tony Award for Best Musical.[^3] In my data tidying, I removed stop words and character names, and I also edited iterations of similar words for clarity of analysis (e.g., all "pies" became "pie").

[^3]: A notable absence is *A Funny Thing Happened on the Way to the Forum*, which also won Best Musical, but I already had an even number of shows and I didn't want to disrupt the balance.

::: callout-note
The custom DSST 389 ChatGPT was utilized to help with my understanding of topic model limitations and most effective uses: https://chatgpt.com/share/6807c556-1fd8-800b-82fc-8ad04479347c
:::

## Data Tidying

```{r}
#| label: get-packages
#| echo: false
#| warning: false
#| results: false

# Download necessary packages

library(tidytext)
library(stringr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(tidymodels)
library(textrecipes)
library(kknn)
library(glmnet)
library(corrplot)
library(parallel)
library(future)
library(textdata)
library(stm)
library(topicmodels)
library(reshape2)
library(ldatuning)
library(tm)
library(glue)

theme_set(theme_minimal())

set.seed(123)

all_cores <- parallel::detectCores()
plan(multisession, workers = all_cores)
```

```{r}
#| label: load-data

# Read in data
merrily <- read.csv("../data/Merrily.csv")

woods <- read.csv("../data/Woods.csv")

sweeney <- read.csv("../data/Sweeney.csv")

company <- read.csv("../data/Company.csv")

sunday <- read.csv("../data/Sunday2.csv")

night <- read.csv("../data/Night.csv")

passion <- read.csv("../data/Passion.csv")

follies <- read.csv("../data/Follies.csv")
```

```{r}
#| label: tidy-data

# Tidy datasets
merrily <- merrily |>
  mutate(linenumber = row_number()) |>
  rename(musical = Musical, song = Song, text = Text)

woods <- woods |>
  mutate(linenumber = row_number()) |>
  rename(musical = Musical, song = Song, text = Text)

sweeney <- sweeney |>
  mutate(linenumber = row_number()) |>
  rename(musical = Musical, song = Song, text = Text) |>
  mutate(song = if_else(song == "", "The Contest", song))

company <- company |>
  mutate(linenumber = row_number()) |>
  rename(musical = Musical, song = Song, text = Text) |>
  mutate(text = str_replace_all(text, "[’]", "'"))

sunday <- sunday |>
  mutate(linenumber = row_number()) |>
  rename(musical = Musical, song = Song, text = Text)

night <- night |>
  mutate(linenumber = row_number()) |>
  mutate(musical = if_else(musical == "", "A Little Night Music", musical))

passion <- passion |>
  mutate(linenumber = row_number())

follies <- follies |>
  mutate(linenumber = row_number())
```

## Data Tokenization

I made a list of names from each show in order to avoid skewing the data based on the "rareness" of the character names. I also created a list of common Sondheim "stop words" like "ah" and "la."

```{r}
#| label: tokenize-data

# Create list of common names for each show
names <- c(
  "charley", "frank", "mary", "beth", "sweeney", "fleet", "pirelli",
  "todd", "beadle", "johanna", "deedle", "toby", "anthony", "turpin",
  "lovett", "robert", "lucy", "bobby", "bubi", "bob", "harry", "amy",
  "paul", "joanne", "robby", "peter", "david", "susan", "sarah", "george",
  "louis", "louis's", "dot", "giorgio", "clara", "fosca", "jessie",
  "buddy", "tony's", "buddy's", "ben", "margie", "sally", "fredrik",
  "henrik", "anne", "petra", "desiree", "charlotte", "signora"
)

# Create list of common Sondheim stop words
sondheim_stop <- c(
  "la", "ah", "bah", "whadaya",
  "ding", "dong", "em", "gonna",
  "shh", "psst", "da", "bzzz"
)

# Tokenize and tidy data sets
merrily_tidy <- merrily |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

woods_tidy <- woods |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

sweeney_tidy <- sweeney |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

company_tidy <- company |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

sunday_tidy <- sunday |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

night_tidy <- night |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

passion_tidy <- passion |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))

follies_tidy <- follies |>
  unnest_tokens(word, text) |>
  anti_join(stop_words, join_by(word))
```

```{r}
#| label: create-corpus

# Create Sondheim corpus
sondheim_corpus <- bind_rows(
  merrily_tidy, woods_tidy,
  sweeney_tidy, company_tidy,
  sunday_tidy, night_tidy,
  passion_tidy, follies_tidy
)

# Edit strings for clearer analysis
sondheim_corpus <- sondheim_corpus |>
  filter(!(word %in% names), !(word %in% sondheim_stop)) |>
  mutate(word = str_replace_all(word, "'s", "")) |>
  mutate(word = str_replace_all(word, "pies", "pie")) |>
  anti_join(stop_words, join_by(word))

# Factor shows for order
sondheim_corpus <- sondheim_corpus |>
  mutate(musical = factor(musical,
    levels = c(
      "Follies", "Passion", "Company",
      "Into the Woods", "A Little Night Music",
      "Merrily We Roll Along",
      "Sunday in the Park with George",
      "Sweeney Todd: The Demon Barber of Fleet Street"
    )
  ))
```

```{r}
#| label: corpus-lengths
#| tbl-cap: "Length (in words) of each Sondheim musical"

# Create list of show lengths
sondheim_lengths <- sondheim_corpus |>
  group_by(musical) |>
  summarize(article_length = n()) |>
  ungroup()

sondheim_lengths |>
  arrange(desc(article_length))
```

## Data Exploration

```{r}
#| label: word-lengths
#| tbl-cap: "Most common words in each Sondheim musical"

# Data exploration: most common words
word_ranks <- sondheim_corpus |>
  count(musical, word, sort = TRUE) |>
  group_by(musical) |>
  mutate(rank = row_number())

word_ranks |>
  group_by(musical) |>
  slice_head(n = 5)
```

```{r}
#| label: tbl-top-words
#| tbl-cap: "Most common words in entire Sondheim corpus"
sondheim_corpus |>
  count(word, sort = TRUE) |>
  slice_head(n = 10)
```

In an interview with *The Paris Review*, Sondheim said, "Two of the hardest words in the language to rhyme are life and love. Of all words!"[^4] Unfortunately for Sondheim, as shown in @tbl-top-words, 'love' and 'life' are both in the top five most-used words of his lyrical corpus. As discussed in the introduction, love is one of Sondheim's most explored themes, so this is not a surprise. Notably, because 'love' and 'life' have high term frequencies, when we compare terms across each show (document) using tf-idf, these terms will have low scores. This will boost terms specific to each corpus to help us showcase how the characters/narratives in each show are unique. After exploring the data with tf-idf, I will use sentiment analysis to help accentuate the differences in sentiment between shows, in order to prepare for interpreting topic models.

[^4]: Lipton, James. 1997. "Stephen Sondheim, The Art of the Musical." *The Paris Review*. https://www.theparisreview.org/interviews/1283/the-art-of-the-musical-stephen-sondheim

## tf-idf

```{r}
#| label: tf-idf-prep

# tf-idf preparation
n_docs <- sondheim_corpus |>
  select(musical) |>
  n_distinct()

idf <- sondheim_corpus |>
  distinct(musical, word) |>
  count(word, name = "word_docs") |>
  mutate(idf = log(n_docs / word_docs))
```

```{r}
#| label: tf-idf-top
#| echo: false
#| results: false

# List of tf-idf by show
sondheim_corpus |>
  count(musical, word) |>
  left_join(sondheim_lengths, join_by(musical)) |>
  mutate(tf = n / article_length) |>
  left_join(idf, join_by(word)) |>
  mutate(tf_idf = tf * idf) |>
  group_by(musical) |>
  arrange(desc(tf_idf)) |>
  slice_head(n = 5)
```

```{r, fig.height = 6, fig.width = 7}
#| label: fig-tf-idf
#| fig-cap:  "tf-idf scores for Sondheim's musicals."

# Top five tf-idf by show
labels <- sondheim_corpus |>
  count(musical, word) |>
  left_join(sondheim_lengths, join_by(musical)) |>
  mutate(tf = n / article_length) |>
  left_join(idf, join_by(word)) |>
  mutate(tf_idf = tf * idf) |>
  group_by(musical) |>
  arrange(desc(tf_idf)) |>
  slice_head(n = 5)

# Visualize tf-idf
sondheim_corpus |>
  count(musical, word) |>
  left_join(sondheim_lengths, join_by(musical)) |>
  mutate(tf = n / article_length) |>
  left_join(idf, join_by(word)) |>
  mutate(tf_idf = tf * idf) |>
  arrange(desc(tf_idf)) |>
  filter(tf_idf > 0) |>
  ggplot(aes(x = musical, y = tf_idf, fill = musical)) +
  geom_violin() +
  geom_label_repel(data = labels, aes(label = word), max.overlaps = 20) +
  scale_y_log10() +
  scale_fill_brewer(palette = "Set2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(override.aes = aes(color = NA))) +
  theme(legend.position = "none") +
  labs(title = "tf-idf of popular Sondheim musicals", x = "Musical")
```

As shown in @fig-tf-idf, the results of the tf-idf point to the highly specific nature of Sondheim's shows (e.g., *Sweeney Todd* and *Into the Woods*). Other shows have intriguing terms with high scores like "perpetual" in *A Little Night Music* or "mutter" in *Merrily We Roll Along*. However, these scores can be attributed to one song in each show skewing the entire corpus evaluation (i.e., "Perpetual Anticipation" and "Franklin Shepard, Inc.," respectively). Next, I'll use sentiment analysis to explore each show's overall sentiment to help provide further context for the topic models I'll create later.

## Sentiment Analysis: afinn

```{r}
#| label: afinn-setup

# Load sentiment lexicon
afinn <- get_sentiments("afinn")

# Compute sentiment scores for each song
sentiment_scores <- sondheim_corpus |>
  left_join(afinn, join_by(word)) |>
  group_by(song, musical) |>
  summarise(
    afinn_net = sum(value, na.rm = TRUE),
    lyric_len = n(),
    afinn_normalized = afinn_net / lyric_len, .groups = "drop"
  ) |>
  ungroup()
```

```{r}
#| label: afinn_scores
#| tbl-cap: "Average sentiment of Sondheim musicals"

# Compute the overall average sentiment
overall_avg_sentiment <- sentiment_scores |>
  summarize(
    avg_sentiment = mean(afinn_normalized, na.rm = TRUE),
    .groups = "drop"
  ) |>
  round(4) |>
  pull(avg_sentiment)

# Get sentiment at Album level
musical_sentiment <- sentiment_scores |>
  group_by(musical) |>
  summarize(
    avg_sentiment = mean(afinn_normalized, na.rm = TRUE),
    .groups = "drop"
  )

musical_sentiment |>
  arrange(desc(avg_sentiment))

sweeney_sentiment <- sentiment_scores |>
  filter(musical == "Sweeney Todd: The Demon Barber of Fleet Street") |>
  summarize(
    avg_sentiment = mean(afinn_normalized, na.rm = TRUE),
    .groups = "drop"
  ) |>
  round(4) |>
  pull(avg_sentiment)

passion_sentiment <- sentiment_scores |>
  filter(musical == "Passion") |>
  summarize(
    avg_sentiment = mean(afinn_normalized, na.rm = TRUE),
    .groups = "drop"
  ) |>
  round(4) |>
  pull(avg_sentiment)


musical_avg_sentiment <- musical_sentiment |>
  pull(avg_sentiment)
```

Sondheim's shows portray a wide variety of sentiment, with his average sentiment landing at `{r} overall_avg_sentiment`. This is somewhat surprising as shows like *Into the Woods* and *Sweeney Todd* are known for their violence and darker themes, yet neither of these shows average in the negatives. Below, in @fig-shows-afinn, the average scores of each Sondheim musical are visualized against the average sentiment of the entire corpus.

```{r}
#| label: fig-shows-afinn
#| fig-cap: "Average sentiment for each Sondheim musical"

# Visualize average sentiment of Sondheim musicals
musical_sentiment |>
  mutate(
    musical = factor(
      musical,
      levels = c(
        "Follies", "Passion", "Company",
        "Into the Woods", "A Little Night Music",
        "Sunday in the Park with George",
        "Sweeney Todd: The Demon Barber of Fleet Street"
      )
    )
  ) |>
  ggplot(aes(x = musical, y = avg_sentiment, fill = musical)) +
  geom_col() +
  geom_hline(
    yintercept = overall_avg_sentiment,
    color = "red", linetype = 2,
    linewidth = 0.8
  ) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Average sentiment of Sondheim musicals",
    x = "Musical",
    y = "Average Sentiment (normalized)"
  )
```

Next, let's run a quick ANOVA test to see if these differences in sentiment are statistically significant:

```{r}
#| label: aov-sentiment

# Run ANOVA on sentiment scores by album
aov_sentiment <- sentiment_scores |>
  aov(afinn_normalized ~ musical, data = _)

summary(aov_sentiment)
```

```{r}
#| label: p-value
#| warning: false
#| echo: false
#| result: false

# Pull p-value from ANOVA output
p_value <- summary(aov_sentiment)[[1]][["Pr(>F)"]][1] |>
  round(3)
```

As shown in the ANOVA output, the differences between average mean sentiment for each Sondheim musical are not statistically significant (p = `{r} p_value`). One thing to consider is that how we perceive the "tone" of a show is heavily influenced by the music and orchestration. *Sweeney Todd* has many songs written in minor keys, whereas shows like *Company* use more major keys. Therefore, even if the lyrics themselves don't differ as much in sentiment, this analysis is unable to account for the influence of the music itself. 

Next, I'll use topic modeling to explore the topic of 'love' in Sondheim's corpus -- first in a global model and then in two individual show models.

## Topic Modeling

```{r}
#| label: sondheim_dtm
#| warning: false

# Set up Sondheim dtm
sondheim_untidy <- bind_rows(
  merrily, woods, sweeney, company,
  sunday, night, passion, follies
)

sondheim_tracks <- sondheim_corpus |>
  select(musical, song) |>
  group_by(musical) |>
  distinct() |>
  mutate(track = row_number())

sondheim_songs <- sondheim_untidy |>
  left_join(sondheim_tracks, join_by(musical, song)) |>
  group_by(musical, song, track) |>
  summarize(full_text = glue_collapse(text, sep = " ")) |>
  ungroup() |>
  mutate(doc_id = row_number()) |>
  arrange(doc_id)

# Tidy new Sondheim corpus
sondheim_words <- sondheim_songs |>
  unnest_tokens(word, full_text) |>
  mutate(word = str_replace_all(word, "'s", "")) |>
  anti_join(stop_words, join_by(word)) |>
  filter(!(word %in% names), !(word %in% sondheim_stop)) |>
  mutate(word = str_replace_all(word, "pies", "pie")) |>
  filter(str_length(word) > 3)

sondheim_dtm <- sondheim_words |>
  count(doc_id, word) |>
  cast_dtm(document = doc_id, term = word, value = n)
```

```{r}
#| label: q-lda-best

# Create lda model
k_best <- 10

lda_best <- LDA(
  sondheim_dtm,
  method = "Gibbs",
  k = k_best,
  control = list(seed = 123)
)
```

```{r}
#| label: fig-visualize-top-terms
#| fig-cap: "Top topics and terms in entire Sondheim corpus"

lda_best |>
  tidy(matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered() +
  labs(title = "Top Terms per Topic (k = 10)")
```

```{r}
#| label: fig-visualize-love
#| fig-cap: "Top terms for topic 'love' in entire Sondheim corpus"

# Create visualization of top terms in topic 'love'
tidy_lda_best <- tidy(lda_best, matrix = "beta")

love_topics <- tidy_lda_best |>
  group_by(topic) |>
  arrange(desc(beta)) |>
  mutate(topic_rank = row_number()) |>
  filter(topic_rank <= 3, term == "love") |>
  pull(topic)

tidy_lda_best |>
  filter(topic %in% love_topics) |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = as_factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered()
```

After exploring between 5 and 15 topics, I decided to choose 10 because it provided the most cohesive overview of 'love' in Sondheim's corpus (in my somewhat qualified opinion). A couple of shows definitely skew the top terms, but this is to be expected due to inequalities in tf of 'love' between shows. As shown in @fig-love-count, *Passion*, *Follies*, and *Company* all contributed the most instances of 'love' to this model. For example, "marry" and "ready" can be attributed to *Company* because grappling with the implications of marriage, loneliness, and love is a central theme of the show.

```{r}
#| label: fig-love-count
#| fig-cap: "Instances of the term 'love' in Sondheim musicals"

sondheim_corpus |>
  group_by(musical) |>
  filter(word == "love") |>
  count(word) |>
  ggplot(aes(x = musical, y = n, fill = musical)) +
  geom_col() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Instances of the term 'love' in Sondheim musicals",
    x = "Musical"
  )
```

```{r}
#| label: love-stats
#| echo: false
#| result: false

passion_love <- sondheim_corpus |>
  filter(musical == "Passion", word == "love") |>
  count(word) |>
  pull(n)

sweeney_love <- sondheim_corpus |>
  filter(
    musical == "Sweeney Todd: The Demon Barber of Fleet Street",
    word == "love"
  ) |>
  count(word) |>
  pull(n)
```

Now, I'll create topic models for two shows that have higher frequencies of 'love' but differ highly in sentiment. *Passion* has the highest frequency of 'love' (n = `{r} passion_love`) and the highest average sentiment (`{r} passion_sentiment`). For the best contrast, I'll compare *Passion* to *Sweeney Todd*, which has (n = `{r} sweeney_love`) and an average sentiment of `{r} sweeney_sentiment`.  

::: callout-note
The k-values for both musicals were chosen through trial-and-error in order to find the most representative list of terms (again, through my somewhat qualified opinion).
:::

### Topic Modeling by Musical: *Passion*

```{r}
#| label: passion_dtm
#| echo: false
#| results: false
#| warning: false

# Creat Passion dtm
passion_songs <- passion |>
  left_join(sondheim_tracks, join_by(musical, song)) |>
  group_by(musical, song, track) |>
  summarize(full_text = glue_collapse(text, sep = " ")) |>
  ungroup() |>
  mutate(doc_id = row_number()) |>
  arrange(doc_id)

passion_words <- passion_songs |>
  unnest_tokens(word, full_text) |>
  anti_join(stop_words, join_by(word)) |>
  filter(!(word %in% names), !(word %in% sondheim_stop)) |>
  mutate(word = str_replace_all(word, "'s", "")) |>
  mutate(word = str_replace_all(word, "pies", "pie")) |>
  filter(str_length(word) > 3)

passion_dtm <- passion_words |>
  count(doc_id, word) |>
  cast_dtm(document = doc_id, term = word, value = n)
```

```{r}
#| label: fig-passion-topics
#| echo: false
#| results: false
#| fig-cap: "Top topics and terms in *Passion*"

# Creat lda model for passion
lda_passion <- LDA(
  passion_dtm,
  method = "Gibbs",
  k = 5,
  control = list(seed = 123)
)

lda_passion |>
  tidy(matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered() +
  labs(title = "Top Terms per Topic (k = 10)")
```

```{r}
#| label: fig-passion-love
#| fig-cap: "Top terms for topic 'love' in *Passion*"

# Create visualization of top terms in topic 'love'
tidy_lda_passion <- tidy(lda_passion, matrix = "beta")

love_passion <- tidy_lda_passion |>
  group_by(topic) |>
  arrange(desc(beta)) |>
  mutate(topic_rank = row_number()) |>
  filter(topic_rank <= 3, term == "love") |>
  pull(topic)

tidy_lda_passion |>
  filter(topic %in% love_passion) |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = as_factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered()
```

As shown in @fig-passion-love, *Passion*'s exploration of love is broader, more 'cliche', and contains terms that one would expect (e.g., "happiness," "mine," "beautiful", etc.). However, there are hints of darker themes through words like "unhappy" and "forgive," suggesting a more complex attitude toward love than apparent at first glance.

### Topic Modeling by Musical: *Sweeney Todd*

```{r}
#| label: sweeney_dtm
#| echo: false
#| results: false
#| warning: false

# Create Sweeney dtm
sweeney_songs <- sweeney |>
  left_join(sondheim_tracks, join_by(musical, song)) |>
  group_by(musical, song, track) |>
  summarize(full_text = glue_collapse(text, sep = " ")) |>
  ungroup() |>
  mutate(doc_id = row_number()) |>
  arrange(doc_id)

sweeney_words <- sweeney_songs |>
  unnest_tokens(word, full_text) |>
  anti_join(stop_words, join_by(word)) |>
  filter(!(word %in% names), !(word %in% sondheim_stop)) |>
  mutate(word = str_replace_all(word, "'s", "")) |>
  mutate(word = str_replace_all(word, "pies", "pie")) |>
  filter(str_length(word) > 3)

sweeney_dtm <- sweeney_words |>
  count(doc_id, word) |>
  cast_dtm(document = doc_id, term = word, value = n)
```

```{r}
#| label: fig-sweeney-topics
#| fig-cap: "Top topics and terms in *Sweeney Todd*"
#| echo: false
#| results: false

# Create lda model for Sweeney
lda_sweeney <- LDA(
  sweeney_dtm,
  method = "Gibbs",
  k = 11,
  control = list(seed = 123)
)

lda_sweeney |>
  tidy(matrix = "beta") |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered() +
  labs(title = "Top Terms per Topic (k = 10)")
```

```{r}
#| label: fig-sweeney-love
#| fig-cap: "Top terms for topic 'love' in *Sweeney Todd*"

# Create visualization of top terms in topic 'love'
tidy_lda_sweeney <- tidy(lda_sweeney, matrix = "beta")

love_sweeney <- tidy_lda_sweeney |>
  group_by(topic) |>
  arrange(desc(beta)) |>
  mutate(topic_rank = row_number()) |>
  filter(topic_rank <= 3, term == "love") |>
  pull(topic)

tidy_lda_sweeney |>
  filter(topic %in% love_sweeney) |>
  group_by(topic) |>
  slice_max(beta, n = 10) |>
  ungroup() |>
  mutate(term = reorder_within(term, beta, topic)) |>
  ggplot(aes(beta, term, fill = as_factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~topic, scales = "free") +
  scale_y_reordered()
```

Contrastly, as shown in @fig-sweeney-love, in *Sweeney Todd*, 'love' is associated with a plethora of terms, obvious and not. For example, "heaven," "wait," and "nice" are fairly innocent, but the topic of love in *Sweeney Todd* becomes more complex when we consider words like "worst," "trouble," and "harm." The presence of these negative words speaks to the pain that Sweeney (and other characters) feel when grappling with love, and Sondheim uses *Sweeney Todd* to explore how love and morality intertwine -- for good and ill.

## Conclusion

As explored above, Sondheim's lyrical corpus is diverse and his characters/narration are often highly specific. Despite this (or perhaps because of this), Sondheim's work brilliantly navigates broad themes of love and life in accessible, meaningful, and heart-wrenching ways. 'Love' emerged as the most common word in Sondheim's corpus, and through tf-idf, sentiment analysis, and topic modeling, I highlighted the varied ways in which Sondheim explores love through unique characters and narration. Shows with a more positive average sentiment (e.g., *Passion*) still approached love with complexity, balancing 'cliche' terms (e.g., "happiness" and "mine") with more brooding ones (e.g., "unhappy" and "forgive"). Similarly, shows with more negative average sentiment (e.g., *Sweeney Todd*) included both hopeful and melancholy terms associated with love (e.g., "heaven" vs. "harm" and "nice" vs. "trouble), illustrating the intricate nature of Sondheim's lyric writing and storytelling. Importantly, as previously mentioned, lyrics only tell part of the story when examining love -- without the context of orchestration, staging, and acting, we cannot explicate the full story of Sondheim's narrative intent. Nonetheless, analyzing his lyrics separately provides important insight into his lyrical genius.
